---
title: "Haverhill_EDA"
author: "Haverhill Jing-Yu Team"
output: pdf_document
---


## The Haverhill dataset
```{r q0, include=FALSE}
library(tidyverse)

```

```{r q1, echo=FALSE}
Haverhill <- read_csv("Haverhill.csv",
                      col_types = cols(
                        Request_ID = col_double(),
                        Master_Request = col_double(),
                        Has_Links = col_logical(),
                        Status_Code = col_double(),
                        Work_Order_Created = col_logical(),
                        Request_Type_ID = col_double(),
                        Request_Type_Private = col_logical(),
                        Department_ID = col_double(),
                        Street_ID = col_double(),
                        City_ID = col_double(),
                        Cross_Street_ID = col_double(),
                        District_ID = col_double(),
                        Escalations = col_double(),
                        Priority = col_double(),
                        Longitude = col_double(),
                        Latitude = col_double(),
                        Origin_ID = col_double(),
                        Ward = col_character(),
                        Ward_Precinct = col_character(),
                        CDBG = col_integer()
                      
                      ))
                     
Haverhill
```


```{r q2.1, echo=FALSE}
library(sf)

pw <- st_read("haverhill_pw/Hav_Precincts_Wards_WGS84.shp", quiet = TRUE)
pw <- pw %>% filter(Map_Year == 2011) %>% st_transform(4326)


```

```{r q2.2, echo=FALSE}
library(sf)

cdbg <- st_read("haverhill_cdbg/Hav_CDBG_Area_WGS84.shp", quiet = TRUE)
cdbg <- cdbg %>% st_transform(4326)


```

```{r q3, echo=FALSE}
requests <- Haverhill %>%
st_as_sf(coords = c("Longitude", "Latitude")) %>%
st_set_crs(4326)
```

## The Heatmap of Requests
```{r q4, echo=FALSE}
ggplot() + 
  geom_sf(data = pw) +
  geom_sf(data = cdbg, color = "red") +
  geom_sf(data = requests, alpha = .01)

 
```


## The Number of Requests per Ward&Precinct
```{r q5, echo=FALSE}
Haverhill %>% count(Ward_Precinct) %>% arrange(desc(n))

Haverhill %>% count(Ward_Precinct) %>%
  ggplot(aes(Ward_Precinct,n)) + geom_bar(stat = "identity")
```


## The Number of Requests in CDBG
```{r q6, echo=FALSE}
Haverhill %>% count(CDBG) %>% arrange(desc(n))

Haverhill %>% count(CDBG) %>%
  ggplot(aes(CDBG,n)) + geom_bar(stat = "identity")

```


## Haverhill Request Type Frequency
```{r q7, echo=FALSE}
Haverhill %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% 
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")


```

## Haverhill CDBG Request Type Frequency
```{r q8, echo=FALSE}
Haverhill %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% 
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")


```

## Haverhill Ward&Precinct 1-1 Request Type Frequency
```{r q1-1, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "1_1") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "1_1") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 1-2 Request Type Frequency
```{r q1-2, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "1_2") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "1_2") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 1-3 Request Type Frequency
```{r q1-3, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "1_3") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "1_3") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 2-1 Request Type Frequency
```{r q2-1, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "2_1") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "2_1") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 2-2 Request Type Frequency
```{r q2-2, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "2_2") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "2_2") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 2-3 Request Type Frequency
```{r q2-3, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "2_3") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "2_3") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 3-1 Request Type Frequency
```{r q3-1, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "3_1") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "3_1") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 3-2 Request Type Frequency
```{r q3-2, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "3_2") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "3_2") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```
 

## Haverhill Ward&Precinct 3-3 Request Type Frequency
```{r q3-3, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "3_3") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "3_3") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 4-1 Request Type Frequency
```{r q4-1, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "4_1") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "4_1") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 4-2 Request Type Frequency
```{r q4-2, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "4_2") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "4_2") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 4-3 Request Type Frequency
```{r q4-3, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "4_3") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "4_3") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 5-1 Request Type Frequency
```{r q5-1, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "5_1") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "5_1") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 5-2 Request Type Frequency
```{r q5-2, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "5_2") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "5_2") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 5-3 Request Type Frequency
```{r q5-3, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "5_3") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "5_3") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 6-1 Request Type Frequency
```{r q6-1, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "6_1") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "6_1") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 6-2 Request Type Frequency
```{r q6-2, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "6_2") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "6_2") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 6-3 Request Type Frequency
```{r q6-3, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "6_3") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "6_3") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 7-1 Request Type Frequency
```{r q7-1, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "7_1") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "7_1") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 7-2 Request Type Frequency
```{r q7-2, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "7_2") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "7_2") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```


## Haverhill Ward&Precinct 7-3 Request Type Frequency
```{r q7-3, echo=FALSE}
Haverhill %>% 
  filter(Ward_Precinct == "7_3") %>% 
  count(Request_Type) %>%
  arrange(desc(n))

Haverhill %>% filter(Ward_Precinct == "7_3") %>%
  count(Request_Type) %>% arrange(desc(n)) %>% head(10) %>%
  ggplot(aes(reorder(Request_Type,n),n)) + 
  geom_bar(stat = "identity") + coord_flip() +
  labs(x = "Request_Type")
  
```

## Top 10 Location Requesting Services
```{r q9, echo=FALSE}
Haverhill %>% count(Address) %>% 
  filter(min_rank(desc(n)) <= 10) %>%
  arrange(desc(n))

```

## Illegal Dumping Sites
```{r q10.1, echo=FALSE}
Haverhill %>% 
  filter(Request_Type == "Trash - Illegal Dumping") %>%
  count(Address) %>% 
  filter(min_rank(desc(n)) <= 10) %>%
  arrange(desc(n))

```

```{r q10.2, echo=FALSE}
illegal_DS <-requests %>% 
  filter(Request_Type == "Trash - Illegal Dumping")

ggplot() + 
  geom_sf(data = pw) +
  geom_sf(data = cdbg, color = "red") +
  geom_sf(data = illegal_DS, alpha = .5)
  

```

## Trash Missed Pickups
```{r q11.1, echo=FALSE}
Haverhill %>% 
  filter(Request_Type == "TRASH - Missed Pick Up") %>%
  count(Address) %>% 
  filter(min_rank(desc(n)) <= 10) %>%
  arrange(desc(n))

```

```{r q11.2, echo=FALSE}
trash_MP <-requests %>% 
  filter(Request_Type == "TRASH - Missed Pick Up")

ggplot() + 
  geom_sf(data = pw) +
  geom_sf(data = cdbg, color = "red") +
  geom_sf(data = trash_MP, alpha = .3)
  

```

## Recycling Missed Pickups
```{r q12.1, echo=FALSE}
Haverhill %>% 
  filter(Request_Type == "RECYCLING - Missed Pick Up") %>%
  count(Address) %>% 
  filter(min_rank(desc(n)) <= 10) %>%
  arrange(desc(n))

```

```{r q12.2, echo=FALSE}
trash_MP <-requests %>% 
  filter(Request_Type == "RECYCLING - Missed Pick Up")

ggplot() + 
  geom_sf(data = pw) +
  geom_sf(data = cdbg, color = "red") +
  geom_sf(data = trash_MP, alpha = .5)
  

```

